---
title: "STAT155-PROJECT3"
author: "Jenny Li, Xiang Li"
date: "2021/12/6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message = FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(broom)
library(knitr)
```

> Introduction to Topic

Sport, as an area receiving increased attention, responds as a result of increased awareness of people to their physical health, especially during the weird year 2020, where the pandemic changes the way of living as well as sport. Some sports benefit from it, some in alternative. Power lifting is a strength sport that consists of three attempts at maximal weight on three lifts: squat, bench press, and dead lift. At the same time, a perfect self-training exercise while staying at home during the pandemic. The central question of power lifting is how to lift successfully. This report contributes to discussing why someone successfully lifted as they expected while others didn’t use logistic regression.

> Research Questions

> Data

> Context

In this report, we’ll use data from the OpenPowerlifting database as of April 2019. OpenPowerlifting is creating a public-domain archive of power lifting history. Power lifting is a sport in which competitors compete to lift the most weight for their class in three separate barbell lifts: the Squat, Bench, and Dead lift. The data is represented by openpowerlifting.csv, which includes all information on the meets (competitions) as well as the competitors who attended those meets worldwide. This dataset tracks 2554800 entries for 668172 lifters from 39909 meets. All the data in the dataset are collected by a team of about eight power lifters from official results posted on federations’ websites. Find more information about the dataset and download the openpowerlifting.csv file on Kaggle: https://www.kaggle.com/open-powerlifting/powerlifting-database. 

```{r load-data, message = FALSE}
# loading dataset
lifting <- read_csv (file = "openpowerlifting.csv")
```

> Cleaning

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Filter data age
lifting1 <- lifting[!is.na(lifting$TotalKg), ]
lifting1 <- lifting1[!is.na(lifting1$BodyweightKg),]
lifting1  %>%
  filter(Age >= 18 & Age <= 40)
#filter the body weight
lifting1 <- lifting[!is.na(lifting$TotalKg),]
lifting1 <- lifting1[!is.na(lifting1$BodyweightKg),]
lifting1 %>%
  filter(BodyweightKg >= 50 & BodyweightKg <= 150)
```

> Linear Regression

> Exploratory Data Analysis

> Model Creation

>Research Question: What is the impact of the body weight of the lifters on the total weight they successfully lifted?

$$[ TotalWeightLifting | bodyweight ] = \beta_0 + \beta_1 bodweight $$

> Fitted Model

> Model Interpretation

> Model Evaluation


> Conclusion 

> General Takeaways

> Limitation

> Appendix
> Works Cited

OpenPowerlifting. “Powerlifting Database.” Kaggle, April 25, 2019. https://www.kaggle.com/open-powerlifting/powerlifting-database.

> R Code
